---
title: "Fitbit Health"
author: "Richard"
date: "12/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages:
```{r}
library(tidyverse)
library(lubridate)
library(skimr)
library(ggplot2)
library(janitor)
library(broom)
library(kableExtra)
```

# 1. Import data

There are 18 data for this project, but we only use necessary files that may be useful for our analysis.
```{r}
Activity <- read.csv("data/raw/csv/dailyActivity_merged.csv")
Sleep <- read.csv("data/raw/csv/sleepDay_merged.csv")
Intensities <- read.csv("data/raw/csv/hourlyIntensities_merged.csv")

calories <- read.csv("data/raw/csv/hourlyCalories_merged.csv")
weight <- read.csv("data/raw/csv/weightLogInfo_merged.csv")
```

```{r}
head(Activity)
Activity = Activity %>% 
  clean_names()
```
```{r}
head(Sleep)
Sleep = Sleep %>% 
  clean_names()
```

```{r}
head(Intensities)
Intensities = Intensities %>% 
  clean_names()
```


# 2. Wrangle and merge data

Change to correct date format:
```{r}
Activity = Activity %>% 
  mutate(activity_date = mdy(activity_date))
head(Activity)
```

```{r}
Sleep = ccccc %>% 
  mutate(
    sleep_day = mdy_hms(sleep_day)
  )
head(Sleep)
```

```{r}
Intensities = Intensities %>% 
  mutate(activity_hour = mdy_hms(activity_hour),
         hour = hour(activity_hour))
```


In the data, `id` is the respondent. Let check how many people in each data:
```{r}
n_distinct(Activity$id) #33
n_distinct(Sleep$id) #24
```
How many unique dates in each file?
```{r}
length(unique(Activity$activity_date)) # 31 days
length(unique(Sleep$sleep_day))
```
Make sure we don't have duplicates in data:
```{r}
duplicated(Activity %>% select(id, activity_date)) %>% sum()
duplicated(Sleep %>% select(id, sleep_day)) %>% sum()
```
Merge data:
```{r}
Df = Activity %>% 
  inner_join(Sleep, by=c("id", "activity_date"="sleep_day"))

head(Df)
```


# 3. Clean up and add data to prepare for analysis




# 4. Descriptive Statistics and Analysis

Summary statistics:
```{r}
skim_without_charts(Df)
```

Some scatter plot of relations between two variables:
```{r}
ggplot(data=Activity, aes(x=total_steps, y=calories)) + 
  geom_point() + geom_smooth() + labs(title="Total Steps vs. Calories")
```


```{r}
ggplot(data=Activity, aes(x=total_distance, y=calories)) + 
  geom_point() + geom_smooth() + 
  labs(title="Total Distance vs. Calories", 
       subtitle = "An inverted relation between total distance and calories")
```

```{r}
f3 = ggplot(data=Activity, aes(x=very_active_distance, y=calories)) + 
  geom_point() + geom_smooth() + 
  labs(
    title="Very active distance vs. Calories",
    subtitle="An inverted U-shaped curve"
  )
f3

ggsave(f3, filename = "fig/f3_very_active_calories.png")

```


```{r}
ggplot(data=Activity, aes(x=moderately_active_distance, y=calories)) + 
  geom_point() + geom_smooth() + 
  labs(title="Moderately active distance vs. Calories")
```
```{r}
f2 = ggplot(data=Activity, aes(x=light_active_distance, y=calories)) + 
  geom_point() + geom_smooth() + 
  theme_minimal() +
  xlab("Light active distance") + ylab("Calories") + 
  labs(
    title="Light active distance vs. Calories",
    subtitle = "Higher light active distance, higher calories"
  )
f2
ggsave(f2, filename = "fig/f2_light_active_calories.png")
```

A regression analysis of active levels and calories:
```{r}
m1 = lm(calories ~ light_active_distance + moderately_active_distance + very_active_distance, data = Activity)
tidy(m1) %>%
  kbl(
    digits = 2,
    caption = "Light active has the largest regression coefficient",
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```


```{r}
ggplot(data=Df, aes(x=calories, y=total_minutes_asleep)) + 
  geom_point() + geom_smooth() + labs(title="Total Steps vs. Total minutes asleep")
```

```{r}
f4 = ggplot(data=Df, aes(x=total_time_in_bed, y=total_minutes_asleep)) + 
  geom_point() + geom_smooth() + 
  labs(
    title="Total time in bed vs. Total minutes asleep"
  )

f4
ggsave(f4, filename = "fig/f4_on_bed_and_asleep.png")

```

Intensities over hours: to see how people's intensities change by hours in a day.
```{r}
Intensities %>% 
  group_by(hour) %>% 
  summarize(
    total_intensity = mean(total_intensity, na.rm=TRUE),
    average_intensity = mean(average_intensity, na.rm=TRUE)
  ) %>% 
  ungroup() %>% 
  ggplot(aes(x=hour, y=total_intensity)) + geom_bar(stat = "identity")
```


Fraction of very active, moderately active, and light active distance: most of time in a day, they spend on light activities.
```{r}
f1 = Activity %>% 
  summarize(
    `very active` = mean(very_active_distance/total_distance*100, na.rm=TRUE),
    moderate = mean(moderately_active_distance/total_distance*100, na.rm=TRUE),
    light = mean(light_active_distance/total_distance*100, na.rm=TRUE),
  ) %>% 
  pivot_longer(everything()) %>% 
  ggplot(aes(x=name, y=value)) + geom_bar(stat = "identity") +
  theme_minimal() +
  xlab("Active levels") + ylab("Percentage") + 
  labs(
    title = "Fraction of distance by active levels",
    subtitle = "Light active attributes for more than 70 percent"
  )

ggsave(f1, filename = "fig/f1_fraction_of_activeness.png")
```

# 5. Visualization











